---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


memoisetools
------------------------------------------------------------------------------

`memoisetools` is a collection of additional caches and helper functions to 
work alongside the [`memoise` package](https://github.com/r-lib/memoise).


This package introduces new caches, new `memoise()` alternatives and functions
for interrogating caches and expiring old objects from a cache.

* New caches:
    * `cache_filesystem2()` - with object timestamping, compression 
      of objects by default, and expiration of results not accessed for a certain time.
    * `cache_memory2()` - with object timestamping, faster `xxhash64`
      used by default, and expiration of results not accessed for a certain time.
* New `memoise::memoise()` alternatives
    * `memoise_with_result_size_limit()` - only store results below a certain 
      size in the cache
    * `memoise_with_mixed_backend()` - have 2 caches in a memoised function, with 
      small results saved in the first cache, and large objects saved in the 
      second cache.
* Helper functions
    * `get_memoise_info()` - to print and return information about the cache(s)
      of a memoised function e.g. how many objects, total size, etc
    * `expire_cache()` - If the cache for a memoised functionhas a timestamp, 
      then this function will deleted cached results older than the specified age


Installation
------------------------------------------------------------------------------

```{r eval=FALSE}
devtools::install_github('coolbutuseless/memoisetools')
```



`get_memoise_info()`
------------------------------------------------------------------------------

`get_memoise_info()` returns information about the caches used by a memoised function.

* `cache` - Cache type. Either 'memory', 'filesystem' or 'gcs or aws'
* Storage location
    * `env` - For 'memory' cache, the R environment in which objects are stored
    * `path` - For 'filesystem' caches, the path to the cache directory
    * `cache_name` - For 'gcs or aws' caches, the name of the cache
* `algo` - Hashing algorithm for creating keys
* `bytes` - Size (in bytes) of the cache
* `n` - Number of objects in the cache
* `has_timestamp` - Does the cache include timestamp information?
* `compress` - Value for 'compress' variable

Note: because memoised functions could have multiple caches (e.g. `memoise_with_mixed_backend`), 
this function returns a list of info for each cache.

```{r}
memoised_rnorm <- memoise::memoise(rnorm)

x <- memoised_rnorm(1000)
y <- memoised_rnorm(12)
z <- memoised_rnorm(1)

memoisetools::get_memoise_info(memoised_rnorm)
```



`cache_filesystem2()`
------------------------------------------------------------------------------

This is a replacement for `memoise::cache_filesystem()` with the following changes:

* use a `tempdir()` if no path specified
* Full absolute path to cache is used, even if initialised with a relative path.  
  This avoids issues as detailed in this [memoise issue on github](https://github.com/r-lib/memoise/issues/51#issuecomment-319993161)
* By default objects saved to filesystem are compressed using gzip compression. 
  (The corresponding [memoise PR on github](https://github.com/r-lib/memoise/pull/70))
* A separate data structure keeps track of the time of all reads/writes to the cache.
* The addition of timestamps allows for expiring objects older than a certain age. 
  See the function `memoisetools::expire_cache()`
* the cache `has_key` method now uses the timestamp cache to determine if a given
  key exists or not. This makes it faster to check if a key exists (as no filesystem
  access is needed), but will cause an error if they file doesn't *actually* exist 
  e.g. if you've deleted the file manually.



`cache_memory2()`
------------------------------------------------------------------------------

This is a replacement for `memoise::cache_memory()` with the following changes:

* A separate data structure keeps track of the time of all reads/writes to the cache.
* The addition of timestamps allows for expiring objects older than a certain age. 
  See the function `memoisetools::expire_cache()`
* Use the faster hash `xxhash64` by default



Expiring objects from the cache
------------------------------------------------------------------------------

With `cache_filesystem2()` and `cache_memory2()`, objects older than a specified
age can be retired from the cache.  I.e. if they have not been read or written 
more recently than the specified time, they will be deleted.

```{r}
memoised_rnorm <- memoise::memoise(rnorm, cache = memoisetools::cache_memory2()) 

memoised_rnorm(1) # stored in cache. 
memoised_rnorm(2) # stored in cache
Sys.sleep(1)      # wait a little bit
memoised_rnorm(1) # recent access to this cached data means it won't be expired

# The following expiry will only delete the cached result for `memoised_rnorm(2)`
# as it has not been read/written in over 1 second
memoisetools::expire_cache(memoised_rnorm, age_in_seconds = 1, verbose = TRUE)

memoised_rnorm(1) # this result is still in the cache
memoised_rnorm(2) # this is a fresh result as the cached version was removed
```



`memoise_with_result_size_limit()`
------------------------------------------------------------------------------

This is a replacement for `memoise::memoise()` which places a limit on how large
an object can be before it is no longer stored in the cache (but simply recalculated
each time).

By default, `memoise::memoise()` will store all results regardless of size.
This works for the majority of cases where you have enough memory and results 
are never too large.

For the  problem I was working on, the function produced many small results and a few very very large results 
(greater than 2GB is size).  If all the big results were cached I'd run out of memory!

In the following example, results over 1000 bytes will not be cached.
```{r collapse=TRUE}
memoised_rnorm <- memoisetools::memoise_with_result_size_limit(rnorm, result_size_limit = 1000)

memoised_rnorm(1) # small enough to cache
memoised_rnorm(1) # getting cached result

head(memoised_rnorm(1000)) # too big to be cached
head(memoised_rnorm(1000)) # so each run produces fresh result
```



`memoise_with_mixed_backend()`
------------------------------------------------------------------------------

This is an adjusted version of `memoise::memoise()` which requires **two** caches
to be set, along with a **size limit**.  Objects smaller than the size limit go
to the first cache, and objects larger than the size limit go to the second cache.

This allows you to cache small results in memory, and send large results to the
filesystem, s3 or google cloud storage.

```{r collapse=TRUE}
memoised_rnorm <- memoisetools::memoise_with_mixed_backend(
  rnorm,
  cache1 = memoisetools::cache_memory2(),
  cache2 = memoisetools::cache_filesystem2(),
  result_size_limit = 1000
)

a <- memoised_rnorm(1) # These 3 results cached to memory
b <- memoised_rnorm(2)
c <- memoised_rnorm(3)

x <- memoised_rnorm(1000)  # These 2 results cached to filesystem
y <- memoised_rnorm(2000)

memoisetools::get_memoise_info(memoised_rnorm)

```
